### 2019-02-21 Datelife_paper1_figures.Rmd
some figures

---
output: pdf_document
fontsize: 12pts
textalign: left  
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
linestretch: 2
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
---

```{r setup, echo = FALSE}
source("Datelife_paper1_global.R")
```

```{r pressure, out.width="100%", include=TRUE, fig.align="center", fig.cap=fig_workflow_main_foot, echo=FALSE}
knitr::include_graphics("fig1.pdf", dpi = 560)
```

```{r eval = TRUE, results = 'asis', out.width="100%", include=TRUE, fig.align="center", fig.cap=fig_runtime_main_foot, echo=FALSE}
download.file(url = "https://github.com/phylotastic/datelife/raw/master/data-raw/benchmark/runtime_tests/2_tests/2_random_spp_names/mdq_plus_gdr_2018.05.29_fill.pdf", destfile = "fig_runtime_main.pdf", mode="wb")
knitr::include_graphics("fig_runtime_main.pdf", dpi = 560)

```