### 2019-05-20 Datelife_paper1_figures.Rmd
figure cap format

---
output: 
  bookdown::pdf_document2:
    keep_tex: yes
textalign: left
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
linestretch: 2
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{caption}
  # - \usepackage{xr}
  #- \externaldocument{name of external document to reference fugures from there in here}
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
source("Datelife_paper1_global.R")
```

<!-- ```{r myfigure, fig.cap = "My figure"} -->
<!-- # fig.show = "hide" AND include = FALSE # does not generate the label for the figure -->
<!-- # fig.cap = " " inactivates the label reference -->
<!-- plot(1:10) -->
<!-- ``` -->

<!-- My plot is called \@ref(fig:myfigure). -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:workflow} -->
<!-- \end{figure} -->
<!-- Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with DateLife, via the R package or through the website. Details on the functions involved on each workflow are shown in `datelife`'s R package vignette. -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:runtime1} -->
<!-- \end{figure} -->
<!-- Computation time of input processing and search across `datelife`s  -->
<!--   chronogram database. -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:schronograms} -->
<!-- \end{figure} -->
<!-- Lineage through time (LTT) plots of Fringillidae (bird family of true finches) source chronograms. The latter are obtained by searching a chronogram database and subsetting chronograms conaning at least two species. Arrows indicate maximum age of each source chronogram and numbers indicate the reference of original studies 1: @barker2012going, 2: @barker2015new, 3: @burns2014phylogenetics, 4: @claramunt2015new, 5: @gibb2015new, 6: @Hedges2015, 7: @hooper2017chromosomal, 8: @Jetz2012, 9: @price2014niche. -->

<!-- \begin{figure}[!ht] -->
<!-- \includegraphics{blank.png} -->
<!-- \caption{} -->
<!-- \label{fig:cvbladj} -->
<!-- \end{figure} -->
<!-- LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarizing information from source chronograms found for the Fringillidae. Arrows indicate maximum age. -->

<!-- \newpage -->

Figure 1. 
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with DateLife, via the R package or through the website. Details on the functions involved on each workflow are shown in `datelife`'s R package vignette.

Figure 2. 
Computation time of input processing and search across `datelife`s chronogram database.

Figure 3. 
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: @barker2012going, 2: @barker2015new, 3: @burns2014phylogenetics, 4: @claramunt2015new, 5: @gibb2015new, 6: @Hedges2015, 7: @hooper2017chromosomal, 8: @Jetz2012, 9: @price2014niche.

Figure 4.
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarizing information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

\newpage

\begin{figure}[!h]
\includegraphics{Fig1.pdf}
\caption{}
\label{fig:workflow}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_runtime1.pdf}
\caption{}
\label{fig:runtime1}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_schronograms1.pdf}
\caption{}
\label{fig:schronograms}
\end{figure}

\newpage

\begin{figure}[!ht]
\includegraphics{fig_crossval_bladj.pdf}
\caption{}
\label{fig:cvbladj}
\end{figure}