### 2019-05-29 Intro.Rmd
intro

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
textalign: left  # this does not work...
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
linestretch: 2
header-includes:
    - \usepackage[left]{lineno}
    - \linenumbers
nocite: |
  @barker2012going, @barker2015new, @burns2014phylogenetics, @claramunt2015new, @gibb2015new, @Hedges2015, @hooper2017chromosomal, @Jetz2012, @price2014niche
---

<!--INTRODUCTION: 1st paragraph: general statement that the field knows to be true and important-->
Time of lineage divergence <!--alt lineages: named biological entities, taxon ages?-->
constitutes a fundamental piece of information for evolutionary
understanding in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000]. This information can be obtained from fossils and from molecular dating studies. The developments in DNA sequencing and phylogenetic dating inferences have encouraged/spurred the application of molecular dating methods to a very large diversity of organisms, greatly increasing the quantity 
<!-- and improving the quality  (do we know this? we would expect thatm but there is no analysis on that really...)-->
of data on taxon ages across the tree of life.
Along with phylogenetic relationships, it allows testing alternative
evolutionary hypothesis in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000].
It has been shown that only a bit of information on time of divergence can improve inferences of phylogenetic distance @Webb2008 bladj
<!--comparative studies, in general, for which knowing
the time context for all life is crucial.-->
Together with geographical distribution patterns, it is a crucial component of historical biogeography studies [@posadas2006historical].
Coupled to species number, it allows studying the tempo and mode of speciation and extinction,
<!-- crucial/vital/critical/essential -->
central for understanding how biodiversity
patterns are shaped across space, time, and clades [@Morlon2014].
<!-- Efforts to have a whole tree of life have been great and here are some examples. -->
<!--INTRODUCTION: 2nd paragraph: important advances in the field-->
In the past two decades, the possibility to obtain good quality DNA sequences
<!-- from a wide variety of organisms became a reality, which,  -->
coupled to methodological developments in phylogenetic and dating inference, allowed
the application of molecular dating methods to a very large diversity
of organisms, greatly increasing the quantity and quality of<!--organisms with--> data on taxon ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of
<!-- both fossil and molecular-based  -->
data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE
and Open Tree of Life (OToL).
<!-- pbdb ??-->
<!--OToL alone hosts now more than 200 chronograms. Cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees) names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
<!-- Methods to include living and fossil lineages are in continued development and increased -->
<!-- usage by the community<!--total evidence and revbayes (fossilized birth-death)-->, -->
<!-- which coupled to better sharing data practices, are greatly contributing to the -->
<!-- accumulation in number and type of available data on taxon ages. -->

<!--INTRODUCTION: 3rd paragraph: problems associated-->
The TimeTree project [@Hedges2006; @Hedges2015; @Kumar2017] has aggregated chronograms
from 3,163 studies, encompassing 97,085 species [@Kumar2017], and continues to grow.
However, even in this gold standard resource, the included taxa only encompass between
0.097 and 3.236% of total species diversity (following taxonomic expert opinion
on the global, extant species numbers, which ranges from 3 to 100 million species
[Mayr2010; Moran2011]). One advantage of TimeTree is that it includes taxa from
across the tree of life, versus more specialized chronogram repositories focusing on land plants
[@stevens2001apw] and mammals [@webb2005phylomatic], birds [@Jetz2012], and fish [@chang2019r]. Users can choose
between a web interface or a mobile app to receive information on divergence times
for the evolutionary history of a lineage, pairs of taxa, all lineages within a
taxon, or a list of taxa. As a science communication tool, TimeTree project is
very powerful: it has a friendly graphical interface, with informative and colorful
outputs, that allows the general public to satisfy curiosity regarding a particular
organism of interest or group of them. It is of limited utility for scientific
studies, however. The thousands of trees that have been entered are unavailable
for examination or reuse; according to the creators (see TimeTree web FAQ), methods
for allowing data downloading have been under discussion for the past several years
yet the primary data remain closed. Moreover, there is no Application Programming
Interface (API) allowing programmatic access to any data, greatly impairing the
possibility of large-scale, automated data-mining, which is not allowed under TimeTree
website's terms of use. The nearly hundred thousand taxon summary chronogram generated
from TimeTree resources is not available with its publication [@Kumar2017] or the
TimeTree website, though the still substantial chronogram from a previous publication
[@Hedges2015] was made available at OToL.
<!--What led to datelife development?
Describe public and research necessities covered by datelife
It is important to use available data on time frame of lineage origin:
To know the state of dating for a group of interest:
What range of estimated ages exist already? Are fossil and molecular
time frames coherent? (e.g., [@Magallon2015c]).
To construct a time tree of life.
For science communication, improve scientific discussions, time-framing
other events of importance in other research areas.
Comparative method: autocorrelation (phylogenetic relatedeness) corrections, phylogenetic signal
-->

<!--INTRODUCTION: 4th paragraph-->
Despite its great importance, analytical tools to summarize available information
on taxon ages for the scientific community are still lacking.
We identified several aspects that might have so far delayed the exploitation of
existing data. First, original chronograms available publicly are scattered across repositories
<!-- (otol tree store, dryad, treebase,  -->
and journals supplementary data, usually with different formats too.
Second, lineage names due to taxonomic idiosincracy can be different among studies
and manual curation of that is usually necessary.
Third, data curation

Platforms that contribute to knowledge in this areas:

Supersmart is an open tool but not easy to use, still requires a lot of curation and knowledge.

Phylogenerator [@pearse2013]

Generate new dates using all available DNA sequence information;
Perform one global analysis using all available information;

Problems or downsides: This might be time consuming for large groups and a lot of
data curation and knowledge on the group of interest is still necessary. For example,
<!--General issues with dating techniques:--> choosing correct fossils for calibration
requires a lot of expertise and knowledge on the group. An incorrect use of fossils
can generate severe bias in dating results [@Sauquet2012c].
<!--Data curation is necessary at some point (is this always true?)-->
Hence, data curation is still an important part of any biological study. The research
community considers it as an important or even crucial step before data analysis.
Hence, automated processes for large data analysis are frequently received with skepticism.

`datelife` palliates this by only using information available from already published
studies, which are ideally constructed using robust information, such as sequence
data and thoughtfully curated fossil calibrations.
<!--Data curation is largely based on taxonomic knowledge; or at least on the interplay between taxonomy and phylogenetics-->
<!--Importance of datelife:-->

<!--INTRODUCTION: 5th paragraph: Goal, objectives and predictions-->
Rapidly increasing data on time of lineage divergence both from molecular and paleontological studies; the increasing importance of use of these data in distant areas of research, often not specialized enough to rapidly obain data on their own; and the lack of an open (both the data sources and the code underlying the analyses) easy to use tool
inspired the development of a prototype `datelife` service over a series
of phylotastic hackathons [@Stoltzfus2013] at the National Evolutionary Synthesis Center.
In this paper we present the first formal description of `datelife`, featuring an improved database of chronograms, more methods to summarize trees, and new functions to visualize data, as well as comparisons of summary trees.
`datelife` is the main service for scaling phylogenetic trees in Phylotastic! system
[@Stoltzfus2013] <!--to construct phylogenetic trees on the fly [http://phylotastic.org/].-->
It can be used through its R package <link to documentation>, and a web interface
(<http://www.datelife.org/query/>)<!-- and an API (still not up, tho)-->.