### 2019-06-01 Datelife_paper1.Rmd
discussion

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
    includes:
      # in_header: preamble.tex
      after_body: [Datelife_paper1_figures_noyaml.tex]
    # fig_caption: no
textalign: left  # this does not work...
geometry: margin = 1in
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
# indent: true # this works
# header-includes:
#    - \setlength\parindent{24pt}  # this works too, but I'm gonna indent manually with 4 spaces
# nope, 4 spaces indents are not recognized, they do something else
linestretch: 2
header-includes:
    # - \usepackage{caption}
    - \usepackage[left]{lineno}
    - \linenumbers
  # - \documentclass[oupdraft]{sysbio} # this is duplicated in tex generated by knitr
  # - \usepackage{xr}
  # - \externaldocument{Datelife_paper1_figures}
    # - \usepackage[utf8]{inputenc}
    # - \usepackage[english]{babel}
    # - \usepackage[document]{ragged2e}
nocite: |
  @barker2012going, @barker2015new, @burns2014phylogenetics, @claramunt2015new, @gibb2015new, @Hedges2015, @hooper2017chromosomal, @Jetz2012, @price2014niche
---

```{r eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, results='hide'}
source("Datelife_paper1_global.R")
```

Running head: DATELIFE: REVEALING THE DATED TREE OF LIFE <!--MAX 50 CHARACTERS, ALL CAPITALS-->

Title: DateLife: Leveraging databases and analytical tools to reveal the dated Tree of Life

Authors: Luna L. Sánchez-Reyes^1^, Brian C. O’Meara^1^

Correspondence address:

1. *Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville, 425 Hesler Biology Building, Knoxville, TN 37996, USA*

Corresponding authors: sanchez.reyes.luna@gmail.com, bomeara@utk.edu

\newpage

**abstract.-**
The combination of new analytical techniques, availability of more fossil and molecular
data, and better practices in data sharing has resulted in a steady accumulation
of chronograms in public and open databases such as Dryad, TreeBASE and Open Tree
of Life, for a large quantity and diversity of organisms. However, this information
remains difficult to synthesize for many biologists and the non-academic community.

**Keywords:** Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations

\newpage
Time of lineage divergence constitutes a fundamental piece of information for evolutionary
understanding in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000], from historical biogeography to species diversification studies [@posadas2006historical; @Morlon2014]. The primary information needed for time of lineage divergence estimation comes from the fossil record. Coupled to molecular phylogenies (and more recenyly to morphological data), the time of divergence of extant and extinct lineages can be reconstructed with molecular dating methods.
Probably encouraged by the great developments in DNA sequencing techniques, phylogenetic inference and molecular dating methods, the number of studies publishing phylogenies with branch lengths proportional to geologic time (hereafter chronograms) have constantly increased in number for the last two decades [@Kumar2017].
Still, generating a chronogram is not an easy task unless you have specialized training. That's why there has been an urge for promoting and facilitating reuse of the vast amount of phylogenetic and time of lineage divergence data that has been generated and made available in publications, for the advantage of research relying on this information [@webb2005phylomatic; @Stoltzfus2013].

Wide interest from the scientific community to make information from phylogenies in general and chronograms  in particular available for consultation and reuse has spurred the creation of public platforms with various goals and characteristics. TreeBASE [@morell1996roots; @Piel2002], the Dryad repository (<http://datadryad.org/>), and the Open Tree of Life [OToL; @Hinchliff2015] platforms store and make available published phylogenies and chronograms for easy scientific reuse. All of them can be queried using automatised web procedures, which permit personalised, large scale queries that are also very fast.
OToL stores trees with branch length information from a wide range of living organisms, implementing a very good metadata structure. Treebase and Dryad repositories also contain trees from all groups of life, but the former does not store branch length information and Dryad stores many other types of biological data, and their metadata does not allow automatic distinction between types of trees, impairing the automatised access to time of lineage divergence information from the latter two databases

Besides keeping a repository to easily store and share expert phylogenetic and chronogram knowledge, OToL also has the primary goal of synthesised all trees in their repository to expose to the community the phylogenetic relationships among all known lineages in a single tree of all life.
All or parts of this synthetic tree can be reused for any purpose; however, it currently only shows tree topology, meaning that it does not expose branch length data of any type, including time of divergence data from their stored chronograms.
Phylomatic [@webb2005phylomatic]
<!-- In a way, phylomatic synthesises trees based on a query. -->
is a platform that synthesizes a tree containing taxa of interest from a user supplied set of trees or from its own database.
It has the advantage that it allows grafting taxa of interest absent from their database, but it is limited to mostly seed plants and mammals. If your query does not match its records, it gives many polytomies back.
The Timetree of Life project focuses on the synthesis of a single chronogram of life [@Hedges2006]. However, the thousands of chronograms compiled for synthesis are only publicly available for visual examination in their website or for download as images, but not for scientific reuse or reanalysis. The latest version of their synthetic chronogram [@Kumar2017] can be queried only through their website in a non-automatised fashion, and only subsets of it can be reused for analyses with the permission of the authors.

Other platforms such as SuperSmart [@antonelli2017supersmart] and phylogenerator [@pearse2013phylogenerator] are focused in automatised _de novo_ chronogram inference, by reusing DNA sequence data to reconstruct phylogenetic trees. However, expert fossil information necessary for subsequent molecular dating analyses still needs to be compiled and curated by the user, rendering them a challenging tool to obtain data on time of lineage divergence for the non-specialist. Also, these tools do not provide information from already existent expert chronograms.

<!-- discussion:
The Angiosperm Phylogeny Website [APW; @stevens2001apw] also specializes in constructing a single synthetic tree, but it focuses in showing phylogenetic relationships between seed plant families only, and it does not store or make available the source trees.

Platforms dedicated to store data on fossils and fossil calibrations: PBDB and The fossil calibration database
can provide data on time of lineage divergence, but it is often not in a phylogenetic context.

To our knowledge, there is no system allowing the comparison of available source information (apw does this a little bit, but it is a verbal comparison more than a graphical or visual comparison). Other types of technical comparisons: which trees are better (which trees represent reality better), but this is of course a very difficult issue.
 -->
A tool for efficient reuse of expert, published data on time of lineage divergence should have an open source database storing chronograms in a format suitable for scientific reuse, an automatised way of accessing the information, and straightforward means of comparing and summarizing chronogram data as needed by the user.
A prototype service aiming to meet this characteristics was developed over a series of hackathons at the National Evolutionary Synthesis Center [@Stoltzfus2013].
In here we present the formal description and implementation of the `datelife` service, constituted by an R package and a web site (<http://www.datelife.org/>). There is still much room for improvement, and flaws and limitations are addressed below. We strived for the current implementation of `datelife` to perform the basic tasks described above, featuring a system for maintenance of an open database of chronograms pulled from public repositories, methods to summarize and compare source chronograms, and new functions to visualize and graphically compare source and summary chronograms. 
<!-- as well as tools for comparison of source and summary trees.  -->
R packages for benchmarking of functionalities and demonstrating services were also developped.

\begin{center}
\textsc{Description}
\end{center}
<!--(BOLD and OToL species names, are they homogeneous/standardized?-->
The basic `datelife` workflow is shown in figure \ref{fig:workflow} and consists of:

1) A user providing at least two taxon names as input, either as tip labels on a tree, or as a simple comma separated character string. The tree can be in newick or phylo format, and can be with or without branch lengths.
2) `datelife` then performs a search across its database of peer reviewed and curated chronograms; identifies and gets source trees with at least two matching input names; drops unmatching taxa from positively identified source trees; and finally transforms each source tree to a patristic matrix named by the citation of the original study. This format facilitates and greatly speeds up all further analyses and summarization algorithms.
3) The user can obtain different types of summaries from the source data including: a) all source chronograms, b) mrca ages of source chronograms, c) citations of studies where source chronograms were originally published, d) a summary table with all of the above, e) a single summary tree of all source chronograms, and f) a report of succesful matches per input taxon name across source chronograms.
4) At this point, users can choose to use all or some source data as calibration points to date a tree of their own making or choosing. <!--, a taxonomic tree-->
5) Users can also simulate age and/or phylogenetic data of input taxa not found in the database. A variety of algorithms are available for this purpose.
6) Finally, users can easily view results graphically as well as construct their own graphs using inbuilt `datelife` graphic generators.

<!--To perform all tasks:-->
To gather, process, and present information, `datelife` builds up from functions
available in several R packages including rotl [@Michonneau2016], ape [@Paradis2004],
geiger [@Harmon2008], paleotree [@Bapst2012a], bold [@Chamberlain2018], phytools [@Revell2012], taxize [@Chamberlain2013; @Chamberlain2018], phyloch [@Heibl2008], phylocomr [@Ooms2018] and rphylotastic [@Omeara2019].

`datelife`'s chronogram database currently contains only chronograms from OToL's
tree repository. Compared to other existing tree repositories, such as TreeBASE and
Dryad, OToL's metadata rich tree store is the only one that supports automatised proper search, identification and handling of chronograms. The TreeBASE repository stores trees without branch lengths, so chronograms stored there do not provide relevant information on time of lineage divergence. The Dryad repository not only stores trees and chronograms, but all sorts of biological data, and their metadata does not allow differentiating between trees with branch lengths proportional to time and any other branch length measure that is not relevant to summarize time of divergence.

`datelife` currently accepts scientific names only. It can be any named clade or binomial specific.
<!--common names (implemented in phylotastic)-->
Chronogram searches are performed at the species level, so when input names correspond
to named clades, `datelife` pulls all accepted species names within the
clade from OToL's reference taxonomy to perform the search.
Currently, searches at the infraspecies level are not allowed, so input names belonging to subspecies or any other infraspecific category are collapsed to the species level.
`datelife` also processes input names with the taxon name resolution service [TNRS; @Boyle2013],
<!--which "corrects misspelled names and authorities, standardizes variant spellings,
and converts nomenclatural synonyms to accepted names" (Boyle et al. 2013)-->
which corrects potentially misspelled names and typos, and standardizes spelling
variations and synonyms <!--to a taxonomic reference of users choosing, default
to otol taxonomy -->, increasing the probability to correctly find the
queried taxa in `datelife`'s chronogram database.

Source chronogram summary trees can be assembled using the Super Distance Matrix
(SDM) supertree construction approach [@Criscuolo2006] or using the median of branch
lengths <!-- and the hierarchical clustering method of ttol-->.
Tree dating and simulation options are performed with various algorithms:
Branch Length Adjuster (BLADJ) is a simple algorithm to distribute ages of undated
nodes evenly, which minimizes age variance in the chronogram [@Webb2008].
PATHd8 is a non-clock, rate-smoothing method [@Britton2007] to date trees.
treePL, is a semi-parametric, rate-smoothing, penalized likelihood dating method
[@Smith2012].
MrBayes [@Huelsenbeck2001; @Ronquist2003] can be used when adding taxa at
random, following a reference taxonomy or a topological constraint. It draws ages
from a pure birth model, as implemented by Jetz and collaborators [-@Jetz2012].
To apply calibrations to a tree, the congruification algorithm described by @Eastman2013
is implemented to find shared nodes between trees (congruent nodes).

<!-- Details on each step are further developed in `datelife`'s R package documentation -->
<!-- `datelife workflow` vignette at (<https://LINK>). -->

\begin{center}
\textsc{Benchmark}
\end{center}
<!--**Benchmark: Testing DateLife computing performance**-->
<!-- Good to test 3 features on a package: performance, speed and scalability-->

`datelife`'s code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor. <!--No multiprocessor tests were performed-->
We registered variation in computing time of query processing and search through the database relative to number of queried taxon names.
Query processing increases roughly linearly with number of input taxon names, and
increases considerably if TNRS service is activated. Up to ten thousand names can be processed and searched in less than 30 minutes. A name search through the chronogram database with an already processed query can be performed in less than a minute, even with a very large number of taxon names (Fig. \ref{fig:runtime1}).
<!--Speed of other datelife workflows? NO, bc it refkects speed from other software mainly. But I will put a sentence about it...
Summarizing `datelife` results processing times: shows sdm speed, clustering methods speed;
generating new trees: shows bladj speed;
Adding dates processing time: reflects more mrbayes speed than anything;
get_bold_otol_tree running time: reflects bold downloading data speed.-->
<!--Speed with different number of chronograms in database? NO-->
<!--Speed of web interface and of r package in computers with different capacities? NOOOO-->
`datelife`'s code performance was evaluated with a set of unit tests designed and
implemented with the R package testthat [@RCoreTeam2018] that were run locally
--using the devtools package [@RCoreTeam2018], and on a public server --via
GitHub, using the continuous integration tool Travis CI (<https://travis-ci.org>). At
present, unit tests cover more than 50% of `datelife`'s code (<https://codecov.io/gh/phylotastic/datelife>). <!-- unit test make code robust, verifiable, debuggable, -->

\begin{center}
\textsc{Example}
\end{center}
<!--**Biological example: Testing DateLife accuracy**-->
<!--Bird (or reptile) chronograms, too long time...
finches is good
nothofagus
Look for all chronograms containing any birds
Or, look for chronograms containing small and old lineages
Determine which clade of birds has the more chronograms (have been dated
more times) and use that as biological example. Fringillidae is a good candidate
Find a clade with at least one chronogram containing all clade's species.
(Penguins look good, but they are giving weird results in SDM)

Remove this chronogram from datelife Results.

Make sdm and median trees and Compare

add taxa with different methods and Compare

Use ltts to compare for now.

think of a test to compare trees, topology- and date-wise

test improvement in synthetic otol tree after using bold sequences. Is it really better resolved? Probably not.
Do not include this analysis in paper, maybe in examples, but i think it is not worthy.
-->
In this section we demonstrate the types of outputs that can be obtained with `datelife`, using as an example the bird family Fringillidae of true finches. We performed a higher-taxon search to obtain all data on lineage divergence available in `datelife`'s database for all recognised species within the Fringillidae (`r length(dq$cleaned_names)` spp. according to the Open Tree of Life taxonomy). There are `r length(dr)` chronograms containing at least two Fringillidae species, published in `r length(unique(names(dr)))` different studies (Fig. \ref{fig:schronograms1}).
<!-- In case we want to cite the studies in here, but they are already cited in the figure caption: [Fig. \ref{fig:schronograms1}; @barker2012going; @barker2015new; @burns2014phylogenetics; @claramunt2015new; @gibb2015new; @Hedges2015; @hooper2017chromosomal; @Jetz2012; @price2014niche]. -->
Data from these source chronograms was used to generate two types of summary chronograms, median and SDM. As explained in the `Description`, data from source chronograms was first summarised into a single distance matrix (using either the median or the SDM method) and then the available node ages were used as calibrations points over a consensus tree topology, to obtain a dated tree with the program BLADJ (Fig. \ref{fig:summaries}). Median summary chronograms are older and have wider variation in maximum ages than chronograms obtained with SDM. In both cases, ages are coherent with source ages.
<!-- (Going from a summary distance matrix Using this method is better than using clustering algorithms  -->
<!-- and hierarchical clustering method which appears to push ages back -->
It is not certain if these chronograms can be used to perform downstream evolutionary analyses. There is currently wide interest in determining this. However, we know that these chronograms are useful for...

Data from source chronograms was also used to date tree topologies with no branch length information and trees with branch lengths in relative substitution rates (Figs. \ref{fig:cvbladj} and \ref{fig:cvbold}). As a form of cross validation, we used tree topologies from each study and calibrated them using information from all other source chronograms. In the absence of branch length data, the ages of internal nodes were approximately recovered in almost all cases (except for studies 3, and 5; Fig. \ref{fig:cvbladj}). Maximum tree ages were only approximately recovered in one case (study 2; Fig. \ref{fig:cvbladj}).
Branch lengths were successfully generated using the BOLD database for all source chronograms. However, dating with PATHd8 (using congruified calibrations) was only successful in
<!-- two studies when expanding calibrations to make them agree (studies 3 and 5), and from  -->
three cases (studies 3, 5, and 9; Fig. \ref{fig:cvbold}). From these, two trees have a different sampling than the original source chronogram, mainly because DNA data for some species is absent from the BOLD. Maximum ages are quite different from source chronograms, but this might be explained also by the differences in sampling between source chronograms and BOLD trees.
More examples and details can be consulted in <https://github.com/LunaSare/datelife_examples>.

<!--Show code here, in the paper? Or show it only in the examples? Or only in the vignette? Write down what about datelife will be shown in each of these.
Paper should show outputs from web site and package.
Vignette will show the R code used to generate the outputs displayed here and in the website.
datelife examples will explain more on what is done in every step.
-->

\begin{center}
\textsc{Flaws, Limitations and Prospects}
\end{center}

The main goal of `datelife` is to make expert information on time of lineage divergence easily accesible for comparison, reuse, and reanalysis, to researchers in all areas of science and with all levels of expertise in the matter. It is a very fast tool that fulfills the quality of openness and does not require any expert biological knowledge from users --besides the names of the organisms they want to work with-- for any of its functionalities. However, it has many flaws. Some of them can be overcome, some of them might represent limitations.

There is only one public tree repository from where `datelife` currently pulls chronograms to construct its database. Other existing tree repositories are not suitable for time of divergence data mining because they either do not store that data or because their metadata does not have enough information to allow automatised chronogram searches. `datelife`'s database has a bit more than 200 chronograms up to the time the manuscript was written. 
<!-- , which probably corresponds to less than 1% (`r 200*100/3987`%) of the information used for TimeTree synthesis  
Even though datelifes chronograms cover this many species from the tree of life compared to this many covered by TimeTree (database or synthetic chronogram?)-->
Incorporation of the many chronograms that are available in data repositories such as Dryad and also from supplementary material of journals to `datelife`'s database is crucial to improve `datelife`'s services. 
Methods to automatically mine Dryad repository for chronogram data could be designed and implemented. 
<!-- What can we do to mine journal databases? -->
<!-- At the same time, the prospect is to keep inviting and doing work encouraging scientists to make their data sets available in OToL's database. -->
At the same time, we would like to increase awereness and emphasize on the importance of data sharing in databases that are maintained by the expert scientific community, requiring expert input and manual curation, such as OToL's tree store.

We introduce different ways of summarizing this data, but are these summary chronograms reliable to study evolutionary patterns? The short answer is that we do not know to what extent the summary trees can be used for reanalyses and in which areas they can be used more reliably than others. Surely they might be useful for ecological analyses, but further studies should be conducted.

As implemented now, we currently by default summarize all source data avilable. Users can subset source data if they have reasons to favor some or one source chronogram over others. Strictly speaking, a good chronogram reflects the real time of divergence accurately and precisely. To our knowledge there is no objective way to determine if an expert chronogram is better than other. Some criteria that have been put forward are the level of sampling and the numer of calibrations used. Scientists usually also favor chronograms coming from studies with primary calibrations to ones from secondary calibrations. `datelife`'s database includes all chronograms derived from expert knowledge, and this includes those derived from secondary calibrations. It has been shown with simulations that in analyses performed with bayesian inference methods, divergence times inferred with secondary calibrations are significantly younger than those inferred with primary calibrations [using secondary calibrations in the same way as primary calibrations are used @schenk2016sec]. But there are different ways to apply secondary calibrations. This information should be included as metadata to facilitate subset of source chronograms. This will need to be manually curated as is done in OToL.

It would be important to evaluate how different source chronograms affect summary chronograms, in order to determine if any type of bias is introduced by certain types of chronograms in particular, so this chronograms are not included in the summary.

Even if all source chronograms have been generated by excellent standards and practices, the evolutionary context they depict might be very different.
The ideal would be to use dates that were obtained from the same DNA context, which should reflect a more homogeneous evolutionary history or hypothesis [@antonelli2017supersmart]. Summarizing trees implies summarizing evolutionary hypothesis.

However, the comparison of source trees in the example above shows that even chronograms obtained with primary data can be very different. There is some evidence that different chronograms will result in very different macroevolutionary histories inferred from them [@title2016macrophylogenies]. In other areas of research, such as ecological and conservation biology, it has been demonstrated that the presence of data on lineage divergence represents an improvement for testing alternative hypothesis, compared to testing in a total absence of such time data. However, we do not know how using wrong or different kinds of age data will affect the outcomes of ecological tests. More work has to be performed in this area.
The comparison of available data has made evident how much discrepancy in time of lineage divergence can be found across expert knowledge. This has been shown for angiosperms previously [@magallon2015metacalibrated], but our examples show that this is a widespread phenomenon that might require further attention.
In the case of the example presented here, most chronograms are new, meaning that we expect that data was generated with high quality and software used was very advanced. However, this is just an assumption. Scores on source data used (quality of alignement: missing data, GC content; tip sampling strategy; phylogenetic inference method; dating inference method; numer of fossils used) But we should also be able to choose on

Improvements:
* fossils as calibrations: Using secondary calibrations can generate biased ages when using bayesian methods, mainly because we don't know what prior to give to secondary calibrations
[@Schenk2016].
* bayesian congruification
* topological congruification

\begin{center}
\textsc{Conclusions}
\end{center}

Time of lineage divergence data are key to many areas of evolutionary studies: trait evolution,
diversification, biogeography, macroecology and more. Obtaining these ages is difficult,
especially for those who want to use phylogenies but who are not systematists, or
do not have the time to acquire and develop the necessary knowledge and data curation skills
to produce chronograms _de novo_. Knowledge on taxon ages is also crucial for
science communication and education.

<!--Mention again uses shown here and outline potential uses not shown here-->
`datelife` allows an easy and fast obtention, as well as comparison of publicly available information
on time of lineage divergence, providing a straightforward way to get an informed idea on the state of knowledge of the timeframe of different regions of the tree of life, allowing identification of regions that need more work or that have conflicting information.
It can also summarize public data, or use it to generate new chronograms based on trees provided by the user.
Both summary and newly generated trees have the potential to be useful to evaluate evolutionary hypothesis in different areas of research. How well this trees work for this purpose still needs to be tested. `datelife` helps with awereness on the variation of expert data, and the actual effect of the underlying data on the results of analyses, nurturing a culture of more cautious interpretation of evolutionary results.

`datelife` is also part of the reproducibility of science movement. It strives to facilitate usage of public data sets as well as encouraging the making of information publicly available. Acknowledging source information and describing workflows and analyses.

it allows exploration of alternative taxonomic and phylogenetic schemes;
it allows rapid exploration of the effect of alternative divergence time hypothesis;
it allows rapid synthesis in a number of different formats; <!-- we should add graphical options too-->

\begin{center}
\textsc{Availability}
\end{center}

`datelife` is free and open source and it can be used through its current website
<http://www.datelife.org/query/>, through its R package, and through Phylotastic's project web portal <http://phylo.cs.nmsu.edu:3000/>.
`datelife`'s website is maintained with RStudio's shiny server and the shiny package open infrastructure, as well as Docker.
`datelife`'s R package stable version is available
for installation from the CRAN repository (<https://cran.r-project.org/package=datelife>)
using the command `install.packages(pkgs = "datelife")` from within R. Development versions
are available from the GitHub repository (<https://github.com/phylotastic/datelife>)
and can be installed using the command `devtools::install_github("phylotastic/datelife")`.

\begin{center}
\textsc{Supplementary Material}
\end{center}

Code used to generate all versions of this manuscript, the biological examples, as well as the benchmark of functionalities can be found in GitHub repositories at <https://github.com/LunaSare/datelife_paper1>, <https://github.com/LunaSare/datelife_examples>, and <https://github.com/LunaSare/datelife_benchmark>, respectively.
<!-- Vignettes
Dryad also?? since it is all in github already, maybe not necessary-->

\begin{center}
\textsc{Funding}
\end{center}

Funding was provided by US National Science Foundation (NSF) grant 1458603

NESCent

Open Tree of Life

University of Tennessee, Knoxville

\begin{center}
\textsc{Acknowledgements}
\end{center}

We thank colleagues from the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used for `datelife`.
The many scientists who publish their chronograms in an open, reusable form, and
the scientists who curate them for deposition in OpenTree.
The NSF for funding nearly all the above, in addition
to the ABI grant that funded this project itself.

\newpage

\begin{center}
\textsc{References}
\end{center}