### 2019-06-01 Datelife_paper1.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin = 1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{}
    \pretitle{\vspace{\droptitle}}
  \posttitle{}
    \author{}
    \preauthor{}\postauthor{}
    \date{}
    \predate{}\postdate{}

\usepackage[left]{lineno}
\linenumbers

\begin{document}

Running head: DATELIFE: REVEALING THE DATED TREE OF LIFE

Title: DateLife: Leveraging databases and analytical tools to reveal the dated Tree of Life

Authors: Luna L. Sánchez-Reyes\textsuperscript{1}, Brian C. O'Meara\textsuperscript{1}

Correspondence address:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \emph{Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville, 425 Hesler Biology Building, Knoxville, TN 37996, USA}
\end{enumerate}

Corresponding authors: \href{mailto:sanchez.reyes.luna@gmail.com}{\nolinkurl{sanchez.reyes.luna@gmail.com}}, \href{mailto:bomeara@utk.edu}{\nolinkurl{bomeara@utk.edu}}

\newpage

\textbf{abstract.-}
The combination of new analytical techniques, availability of more fossil and molecular
data, and better practices in data sharing has resulted in a steady accumulation
of chronograms in public and open databases such as Dryad, TreeBASE and Open Tree
of Life, for a large quantity and diversity of organisms. However, this information
remains difficult to synthesize for many biologists and the non-academic community.

\textbf{Keywords:} Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations

\newpage

Time of lineage divergence constitutes a fundamental piece of information for evolutionary
understanding in many areas of research, from developmental to conservation biology (Felsenstein \protect\hyperlink{ref-Felsenstein1985a}{1985}; Webb \protect\hyperlink{ref-Webb2000}{2000}), from historical biogeography to species diversification studies (Posadas et al. \protect\hyperlink{ref-posadas2006historical}{2006}; Morlon \protect\hyperlink{ref-Morlon2014}{2014}). The primary information needed for time of lineage divergence estimation comes from the fossil record. Coupled to molecular phylogenies, the time of divergence of extant lineages is reconstructed with molecular dating methods.

Probably encouraged by the great developments in DNA sequencing techniques, phylogenetic inference and molecular dating methods, the number of studies publishing phylogenies with branch lengths proportional to geologic time (hereafter chronograms) have constantly increased in number for the last two decades (Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}).
Still, generating a chronogram is not an easy task unless you have specialized training. That's why there has been an urge for promoting and facilitating reuse of the vast amount and still accumulating information that is already available, for the advantage of researchers not specialized in this area (Webb and Donoghue \protect\hyperlink{ref-webb2005phylomatic}{2005}; Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013}).

Wide interest from the scientific community to make information from phylogenies in general and chronograms in particular available for consultation and reuse has spurred the creation of public platforms with various goals and characteristics. TreeBASE (Morell \protect\hyperlink{ref-morell1996roots}{1996}; Piel et al. \protect\hyperlink{ref-Piel2002}{2002}), the Dryad repository (\url{http://datadryad.org/}), and the Open Tree of Life (OToL; Hinchliff et al. \protect\hyperlink{ref-Hinchliff2015}{2015}) platforms are committed to storing and making available published phylogenies and chronograms for easy scientific reuse. All of them can be queried using automatised web procedures, which permit personalised, large scale queries that are also very fast.
OToL's stores trees with branch lengths from a wide range of living organisms, using a very good metadata structure. Treebase and Dryad repositories also contain trees from all groups of life, but the former does not store branch length information and Dryad stores many other types of biological data, and their metadata does not allow automatic distinction between types of trees, impairing the automatised access to time of lineage divergence information.

Besides storing public phylogenetic knowledge, OToL also has the primary goal of exposing a single tree of all life synthesised from trees in their repository, to show phylogenetic relationships of all known lineages.
All or parts of OToL's synthetic tree can be reused for any purpose; however, it currently only focus on tree topology, meaning that it does not incorporate information on time of divergence from their stored chronograms.
Phylomatic (Webb and Donoghue \protect\hyperlink{ref-webb2005phylomatic}{2005})
is a platform that synthesizes a tree containing taxa of interest from a user supplied set of trees or from its own database.
It has the advantage that it allows grafting taxa of interest absent from their database, but it is limited to mostly seed plants and mammals. If your query does not match its records, it gives many polytomies back.
The Timetree of Life project focuses on the synthesis of a single chronogram of life (Hedges et al. \protect\hyperlink{ref-Hedges2006}{2006}). However, the thousands of chronograms compiled for synthesis are only publicly available for visual examination in their website or for download as images, but not for scientific reuse or reanalysis. The latest version of their synthetic chronogram (Kumar et al. \protect\hyperlink{ref-Kumar2017}{2017}) can be queried only through their website in a non-automatised fashion, and only subsets of it can be reused for analyses with the permission of the authors.

Other platforms such as SuperSmart (Antonelli et al. \protect\hyperlink{ref-antonelli2017supersmart}{2017}) and phylogenerator (Pearse and Purvis \protect\hyperlink{ref-pearse2013phylogenerator}{2013}) are focused in automatised \emph{de novo} chronogram inference, by reusing DNA sequence data to reconstruct phylogenetic trees. However, expert fossil information necessary for subsequent molecular dating analyses still needs to be compiled and curated by the user, rendering them a challenging tool to obtain data on time of lineage divergence for the non-specialist. Also, these tools do not provide information from already existent expert chronograms.

A tool for efficient reuse of expert, published data on time of lineage divergence should have an open source database storing data in a format suitable for scientific reuse, an automatised way of accessing the information, and straightforward means of comparing and summarizing the source data as needed by the user.
A prototype service aiming to meet this characteristics was initially developed over a series of hackathons at the National Evolutionary Synthesis Center (Stoltzfus et al. \protect\hyperlink{ref-Stoltzfus2013}{2013}).
In here we present the first formal description and implementation of the \texttt{datelife} service, constituted by and R package and a web site (\url{http://www.datelife.org/}). There is still much room for improvement, and flaws and limitations are addressed below. We strived for the current implementation of \texttt{datelife} to perform the basic tasks described above, featuring a system for maintenance of an open database of chronograms pulled from public repositories, methods to summarize and compare source chronograms, new functions to visualize and graphically compare source and summary chronograms, as well as tools for comparison of source and summary trees. R packages for benchmarking of functionalities and exemplifying services were also developped.

\begin{center}
\textsc{Description}
\end{center}

The basic \texttt{datelife} workflow is shown in figure \ref{fig:workflow} and consists of:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  A user providing at least two taxon names as input, either as tip labels on a tree, or as a simple comma separated character string. The tree can be in newick or phylo format, and can be with or without branch lengths.
\item
  \texttt{datelife} then performs a search across its database of peer reviewed and curated chronograms; identifies and gets source trees with at least two matching input names; drops unmatching taxa from positively identified source trees; and finally transforms each source tree to a patristic matrix named by the citation of the original study. This format facilitates and greatly speeds up all further analyses and summarization algorithms.
\item
  The user can obtain different types of summaries from the source data including: a) all source chronograms, b) mrca ages of source chronograms, c) citations of studies where source chronograms were originally published, d) a summary table with all of the above, e) a single summary tree of all source chronograms, and f) a report of succesful matches per input taxon name across source chronograms.
\item
  At this point, users can choose to use all or some source data as calibration points to date a tree of their own making or choosing. 
\item
  Users can also simulate age and/or phylogenetic data of input taxa not found in the database. A variety of algorithms are available for this purpose.
\item
  Finally, users can easily view results graphically as well as construct their own graphs using inbuilt \texttt{datelife} graphic generators.
\end{enumerate}

\texttt{datelife}'s chronogram database is currently built from OToL's
tree repository. Among currently existing repositories (i.e., TreeBASE,
Dryad), OToL's metadata rich tree store is the only one meeting the requirements
for automatized handling of source trees. TreeBASE trees are stored without branch lengths. Dryad metadata does not allow differentiating between different types of data sets, and trees would need to be manually curated anyways.
\texttt{datelife} currently accepts scientific names only. It can be any named clade or binomial specific.
Chronogram searches are performed at the species level, so when input names correspond
to named clades, \texttt{datelife} pulls all accepted species names within the
clade from OToL's reference taxonomy to perform the search.
Currently, searches at the infraspecies level are not allowed, so input names belonging to subspecies or any other infraspecific category are collapsed to the species level.
\texttt{datelife} also processes input names with the taxon name resolution service (TNRS; Boyle et al. \protect\hyperlink{ref-Boyle2013}{2013}),
which corrects potentially misspelled names and typos, and standardizes spelling
variations and synonyms , increasing the probability to correctly find the
queried taxa in \texttt{datelife}'s chronogram database.

Source chronogram summary tree can be assembled using the Super Distance Matrix
(SDM) supertree construction approach (Criscuolo et al. \protect\hyperlink{ref-Criscuolo2006}{2006}) or using the median of branch
lengths .
Tree dating and simulation options are performed with various algorithms:
Branch Length Adjuster (BLADJ) is a simple algorithm to distribute ages of undated
nodes evenly, which minimizes age variance in the chronogram (Webb et al. \protect\hyperlink{ref-Webb2008}{2008}).
PATHd8 is a non-clock, rate-smoothing method (Britton et al. \protect\hyperlink{ref-Britton2007}{2007}) to date trees.
treePL, is a semi-parametric, rate-smoothing, penalized likelihood dating method
(Smith and O'Meara \protect\hyperlink{ref-Smith2012}{2012}).
MrBayes (Huelsenbeck and Ronquist \protect\hyperlink{ref-Huelsenbeck2001}{2001}; Ronquist and Huelsenbeck \protect\hyperlink{ref-Ronquist2003}{2003}) can be used when adding taxa at
random, following a reference taxonomy or a topological constraint. It draws ages
from a pure birth model, as implemented by Jetz and collaborators (\protect\hyperlink{ref-Jetz2012}{2012}).
To apply calibrations to a tree, the congruification algorithm described by Eastman et al. (\protect\hyperlink{ref-Eastman2013}{2013})
is implemented to find shared nodes between trees (congruent nodes).

To gather, process, and present information, \texttt{datelife} builds up from functions
available in several R packages including rotl (Michonneau et al. \protect\hyperlink{ref-Michonneau2016}{2016}), ape (Paradis et al. \protect\hyperlink{ref-Paradis2004}{2004}),
geiger (Harmon et al. \protect\hyperlink{ref-Harmon2008}{2008}), paleotree (Bapst \protect\hyperlink{ref-Bapst2012a}{2012}), bold (Chamberlain \protect\hyperlink{ref-Chamberlain2018}{2018}), phytools (Revell \protect\hyperlink{ref-Revell2012}{2012}),
taxize (Chamberlain and Szöcs \protect\hyperlink{ref-Chamberlain2013}{2013}; Chamberlain \protect\hyperlink{ref-Chamberlain2018}{2018}), phyloch (Heibl \protect\hyperlink{ref-Heibl2008}{2008}), phylocomr (Ooms and Chamberlain \protect\hyperlink{ref-Ooms2018}{2018})
and rphylotastic (O'Meara et al. \protect\hyperlink{ref-Omeara2019}{2019}).

\begin{center}
\textsc{Benchmark}
\end{center}

\texttt{datelife}'s code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor.
We registered variation in computing time of query processing and search through the database relative to number of queried taxon names.
Query processing increases roughly linearly with number of input taxon names, and
increases considerably if TNRS service is activated. Up to ten thousand names can be processed and searched in less than 30 minutes. A name search through the chronogram database with an already processed query can be performed in less than a minute, even with a very large number of taxon names (Fig. \ref{fig:runtime1}).
\texttt{datelife}'s code performance was evaluated with a set of unit tests designed and
implemented with the R package testthat (R Core Team \protect\hyperlink{ref-RCoreTeam2018}{2018}) that were run locally
--using the devtools package (R Core Team \protect\hyperlink{ref-RCoreTeam2018}{2018}), and on a public server --via
GitHub, using the continuous integration tool Travis CI (\url{https://travis-ci.org}). At
present, unit tests cover more than 50\% of \texttt{datelife}'s code (\url{https://codecov.io/gh/phylotastic/datelife}).

\begin{center}
\textsc{Example}
\end{center}

In this section we demonstrate the types of outputs that can be obtained with \texttt{datelife}, using as an example the bird family Fringillidae of true finches. We performed a higher-taxon search to obtain all data on lineage divergence available in \texttt{datelife}'s database for all recognised species within the Fringillidae (475 spp. according to the Open Tree of Life taxonomy). There are 13 chronograms containing at least two Fringillidae species, published in 9 different studies (Fig. \ref{fig:schronograms1}).
Data from these source chronograms was used to generate two types of summary chronograms, median and SDM. As explained in the \texttt{Description}, data from source chronograms was first summarised into a single distance matrix (using either the median or the SDM method) and then the available node ages were used as calibrations points over a consensus tree topology, to obtain a dated tree with the program BLADJ (Fig. \ref{fig:summaries}). Median summary chronograms are older and have wider variation in maximum ages than chronograms obtained with SDM. In both cases, ages are coherent with source ages.
It is not certain if these chronograms can be used to perform downstream evolutionary analyses. There is currently wide interest in determining this. However, we know that these chronograms are useful for\ldots{}

Data from source chronograms was also used to date tree topologies with no branch length information and trees with branch lengths in relative substitution rates (Figs. \ref{fig:cvbladj} and \ref{fig:cvbold}). As a form of cross validation, we used tree topologies from each study and calibrated them using information from all other source chronograms. In the absence of branch length data, the ages of internal nodes were approximately recovered in almost all cases (except for studies 3, and 5; Fig. \ref{fig:cvbladj}). Maximum tree ages were only approximately recovered in one case (study 2; Fig. \ref{fig:cvbladj}).
Branch lengths were successfully generated using the BOLD database for all source chronograms. However, dating with PATHd8 (using congruified calibrations) was only successful in
three cases (studies 3, 5, and 9; Fig. \ref{fig:cvbold}). From these, two trees have a different sampling than the original source chronogram, mainly because DNA data for some species is absent from the BOLD. Maximum ages are quite different from source chronograms, but this might be explained also by the differences in sampling between source chronograms and BOLD trees.
More examples and details can be consulted in \url{https://github.com/LunaSare/datelife_examples}.

\begin{center}
\textsc{Flaws, Limitations and Prospects}
\end{center}

The main goal of \texttt{datelife} is to make expert information on time of lineage divergence easily accesible for comparison, reuse, and reanalysis, to researchers in all areas of science and with all levels of expertise in the matter. It is a very fast tool that fulfills the quality of openness and does not require any expert biological knowledge from users --besides the names of the organisms they want to work with-- for any of its functionalities. However, it has many flaws. Some of them can be overcome, some of them might represent limitations.

There is only one public repository that has the metadata that datelife needs to perform its tasks and services.
Other databases are not suitable for data mining because of their metadata structure.
OToL's tree database is very large for phylogenetic trees, but only stores a little bit more than 200 chronograms. This database is maintained by the scientific community, so the prospect is to keep inviting and doing work encouraging scientists to make their data sets available in this database.

We introduce different ways of summarizing this data, but are these summary chronograms reliable to study evolutionary patterns? The short answer is that we do not know to what extent the summary trees can be used for reanalyses and in which areas they can be used more reliably than others. Surely they might be useful for ecological analyses, but further studies should be conducted.

As implemented now, we currently by default summarize all source data avilable. Users can subset source data if they have reasons to favor some or one source chronogram over others. To our knowledge there is no objective way to determine if an expert chronogram is better than other. Some criteria that have been put forward are the level of sampling and the numer of calibrations used. Scientists usually also favor chronograms coming from studies with primary calibrations to ones from secondary calibrations.
It would be important to evaluate how different source chronograms affect summary chronograms, in order to determine if any type of bias is introduced by certain types of chronograms in particular, so this chronograms are not included in the summary.
A first step in that direction would be to add data on chronograms that come from primary or secondary calibrations, and maybe the number of calibrations used. This will need to be manually curated.

Even if all source chronograms have been generated by excellent standards and practices, the evolutionary context they depict might be very different.
The ideal would be to use dates that were obtained from the same DNA context, which should reflect a more homogeneous evolutionary history or hypothesis (Antonelli et al. \protect\hyperlink{ref-antonelli2017supersmart}{2017}). Summarizing trees implies summarizing evolutionary hypothesis.

However, the comparison of source trees in the example above shows that even chronograms obtained with primary data can be very different. There is some evidence that different chronograms will result in very different macroevolutionary histories inferred from them (Title and Rabosky \protect\hyperlink{ref-title2016macrophylogenies}{2016}). In other areas of research, such as ecological and conservation biology, it has been demonstrated that the presence of data on lineage divergence represents an improvement for testing alternative hypothesis, compared to testing in a total absence of such time data. However, we do not know how using wrong or different kinds of age data will affect the outcomes of ecological tests. More work has to be performed in this area.
The comparison of available data has made evident how much discrepancy in time of lineage divergence can be found across expert knowledge. This has been shown for angiosperms previously (Magallón et al. \protect\hyperlink{ref-magallon2015metacalibrated}{2015}), but our examples show that this is a widespread phenomenon that might require further attention.
In the case of the example presented here, most chronograms are new, meaning that we expect that data was generated with high quality and software used was very advanced. However, this is just an assumption. Scores on source data used (quality of alignement: missing data, GC content; tip sampling strategy; phylogenetic inference method; dating inference method; numer of fossils used) But we should also be able to choose on

Improvements:
* fossils as calibrations: Using secondary calibrations can generate biased ages when using bayesian methods, mainly because we don't know what prior to give to secondary calibrations
(Schenk \protect\hyperlink{ref-Schenk2016}{2016}).
* bayesian congruification
* topological congruification

\begin{center}
\textsc{Conclusions}
\end{center}

Time of lineage divergence data are key to many areas of evolutionary studies: trait evolution,
diversification, biogeography, macroecology and more. Obtaining these ages is difficult,
especially for those who want to use phylogenies but who are not systematists, or
do not have the time to acquire and develop the necessary knowledge and data curation skills
to produce chronograms \emph{de novo}. Knowledge on taxon ages is also crucial for
science communication and education.

\texttt{datelife} allows an easy and fast obtention, as well as comparison of publicly available information
on time of lineage divergence, providing a straightforward way to get an informed idea on the state of knowledge of the timeframe of different regions of the tree of life, allowing identification of regions that need more work or that have conflicting information.
It can also summarize public data, or use it to generate new chronograms based on trees provided by the user.
Both summary and newly generated trees have the potential to be useful to evaluate evolutionary hypothesis in different areas of research. How well this trees work for this purpose still needs to be tested. \texttt{datelife} helps with awereness on the variation of expert data, and the actual effect of the underlying data on the results of analyses, nurturing a culture of more cautious interpretation of evolutionary results.

\texttt{datelife} is also part of the reproducibility of science movement. It strives to facilitate usage of public data sets as well as encouraging the making of information publicly available. Acknowledging source information and describing workflows and analyses.

it allows exploration of alternative taxonomic and phylogenetic schemes;
it allows rapid exploration of the effect of alternative divergence time hypothesis;
it allows rapid synthesis in a number of different formats;

\begin{center}
\textsc{Availability}
\end{center}

\texttt{datelife} is free and open source and it can be used through its current website
\url{http://www.datelife.org/query/}, through its R package, and through Phylotastic's project web portal \url{http://phylo.cs.nmsu.edu:3000/}.
\texttt{datelife}'s website is maintained with RStudio's shiny server and the shiny package open infrastructure, as well as Docker.
\texttt{datelife}'s R package stable version is available
for installation from the CRAN repository (\url{https://cran.r-project.org/package=datelife})
using the command \texttt{install.packages(pkgs\ =\ "datelife")} from within R. Development versions
are available from the GitHub repository (\url{https://github.com/phylotastic/datelife})
and can be installed using the command \texttt{devtools::install\_github("phylotastic/datelife")}.

\begin{center}
\textsc{Supplementary Material}
\end{center}

Code used to generate all versions of this manuscript, the biological examples, as well as the benchmark of functionalities can be found in GitHub repositories at \url{https://github.com/LunaSare/datelife_paper1}, \url{https://github.com/LunaSare/datelife_examples}, and \url{https://github.com/LunaSare/datelife_benchmark}, respectively.

\begin{center}
\textsc{Funding}
\end{center}

Funding was provided by US National Science Foundation (NSF) grant 1458603

NESCent

Open Tree of Life

University of Tennessee, Knoxville

\begin{center}
\textsc{Acknowledgements}
\end{center}

We thank colleagues from the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used for \texttt{datelife}.
The many scientists who publish their chronograms in an open, reusable form, and
the scientists who curate them for deposition in OpenTree.
The NSF for funding nearly all the above, in addition
to the ABI grant that funded this project itself.

\newpage

\begin{center}
\textsc{References}
\end{center}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-antonelli2017supersmart}{}%
Antonelli A., Hettling H., Condamine F.L., Vos K., Nilsson R.H., Sanderson M.J., Sauquet H., Scharn R., Silvestro D., Töpel M., Bacon C.D., Oxelman B., Vos R.A. 2017. Toward a self-updating platform for estimating rates of speciation and migration, ages, and relationships of Taxa. Systematic Biology. 66:153--166.

\leavevmode\hypertarget{ref-Bapst2012a}{}%
Bapst D.W. 2012. Paleotree: An R package for paleontological and phylogenetic analyses of evolution. Methods in Ecology and Evolution. 3:803--807.

\leavevmode\hypertarget{ref-barker2012going}{}%
Barker F.K., Burns K.J., Klicka J., Lanyon S.M., Lovette I.J. 2012. Going to extremes: Contrasting rates of diversification in a recent radiation of new world passerine birds. Systematic biology. 62:298--320.

\leavevmode\hypertarget{ref-barker2015new}{}%
Barker F.K., Burns K.J., Klicka J., Lanyon S.M., Lovette I.J. 2015. New insights into new world biogeography: An integrated view from the phylogeny of blackbirds, cardinals, sparrows, tanagers, warblers, and allies. The Auk: Ornithological Advances. 132:333--348.

\leavevmode\hypertarget{ref-Boyle2013}{}%
Boyle B., Hopkins N., Lu Z., Raygoza Garay J.A., Mozzherin D., Rees T., Matasci N., Narro M.L., Piel W.H., Mckay S.J., Lowry S., Freeland C., Peet R.K., Enquist B.J. 2013. The taxonomic name resolution service: An online tool for automated standardization of plant names. BMC Bioinformatics. 14.

\leavevmode\hypertarget{ref-Britton2007}{}%
Britton T., Anderson C.L., Jacquet D., Lundqvist S., Bremer K. 2007. Estimating Divergence Times in Large Phylogenetic Trees. Systematic Biology. 56:741--752.

\leavevmode\hypertarget{ref-burns2014phylogenetics}{}%
Burns K.J., Shultz A.J., Title P.O., Mason N.A., Barker F.K., Klicka J., Lanyon S.M., Lovette I.J. 2014. Phylogenetics and diversification of tanagers (passeriformes: Thraupidae), the largest radiation of neotropical songbirds. Molecular Phylogenetics and Evolution. 75:41--77.

\leavevmode\hypertarget{ref-Chamberlain2018}{}%
Chamberlain S. 2018. bold: Interface to Bold Systems API..

\leavevmode\hypertarget{ref-Chamberlain2013}{}%
Chamberlain S.A., Szöcs E. 2013. taxize : taxonomic search and retrieval in R {[}version 2; referees: 3 approved{]}. F1000Research. 2:1--29.

\leavevmode\hypertarget{ref-claramunt2015new}{}%
Claramunt S., Cracraft J. 2015. A new time tree reveals earth history's imprint on the evolution of modern birds. Science advances. 1:e1501005.

\leavevmode\hypertarget{ref-Criscuolo2006}{}%
Criscuolo A., Berry V., Douzery E.J., Gascuel O. 2006. SDM: A fast distance-based approach for (super)tree building in phylogenomics. Systematic Biology. 55:740--755.

\leavevmode\hypertarget{ref-Eastman2013}{}%
Eastman J.M., Harmon L.J., Tank D.C. 2013. Congruification: Support for time scaling large phylogenetic trees. Methods in Ecology and Evolution. 4:688--691.

\leavevmode\hypertarget{ref-Felsenstein1985a}{}%
Felsenstein J. 1985. Phylogenies and the Comparative Method. The American Naturalist. 125:1--15.

\leavevmode\hypertarget{ref-gibb2015new}{}%
Gibb G.C., England R., Hartig G., McLenachan P.A., Taylor Smith B.L., McComish B.J., Cooper A., Penny D. 2015. New zealand passerines help clarify the diversification of major songbird lineages during the oligocene. Genome biology and evolution. 7:2983--2995.

\leavevmode\hypertarget{ref-Harmon2008}{}%
Harmon L., Weir J., Brock C., Glor R., Challenger W. 2008. GEIGER: investigating evolutionary radiations. Bioinformatics. 24:129--131.

\leavevmode\hypertarget{ref-Hedges2006}{}%
Hedges S.B., Dudley J., Kumar S. 2006. TimeTree: A public knowledge-base of divergence times among organisms. Bioinformatics. 22:2971--2972.

\leavevmode\hypertarget{ref-Hedges2015}{}%
Hedges S.B., Marin J., Suleski M., Paymer M., Kumar S. 2015. Tree of life reveals clock-like speciation and diversification. Molecular Biology and Evolution. 32:835--845.

\leavevmode\hypertarget{ref-Heibl2008}{}%
Heibl C. 2008. PHYLOCH: R language tree plotting tools and interfaces to diverse phylogenetic software packages..

\leavevmode\hypertarget{ref-Hinchliff2015}{}%
Hinchliff C.E., Smith S.A., Allman J.F., Burleigh J.G., Chaudhary R., Coghill L.M., Crandall K.A., Deng J., Drew B.T., Gazis R., Gude K., Hibbett D.S., Katz L.A., Laughinghouse H.D., McTavish E.J., Midford P.E., Owen C.L., Ree R.H., Rees J.A., Soltis D.E., Williams T., Cranston K.A. 2015. Synthesis of phylogeny and taxonomy into a comprehensive tree of life. Proceedings of the National Academy of Sciences. 112:12764--12769.

\leavevmode\hypertarget{ref-hooper2017chromosomal}{}%
Hooper D.M., Price T.D. 2017. Chromosomal inversion differences correlate with range overlap in passerine birds. Nature ecology \& evolution. 1:1526.

\leavevmode\hypertarget{ref-Huelsenbeck2001}{}%
Huelsenbeck J.P., Ronquist F. 2001. MRBAYES: Bayesian inference of phylogenetic trees. Bioinformatics. 17:754--755.

\leavevmode\hypertarget{ref-Jetz2012}{}%
Jetz W., Thomas G., Joy J.J., Hartmann K., Mooers A. 2012. The global diversity of birds in space and time. Nature. 491:444--448.

\leavevmode\hypertarget{ref-Kumar2017}{}%
Kumar S., Stecher G., Suleski M., Hedges S.B. 2017. TimeTree: A Resource for Timelines, Timetrees, and Divergence Times. Molecular biology and evolution. 34:1812--1819.

\leavevmode\hypertarget{ref-magallon2015metacalibrated}{}%
Magallón S., Gómez-Acevedo S., Sánchez-Reyes L.L., Hernández-Hernández T. 2015. A metacalibrated time-tree documents the early rise of flowering plant phylogenetic diversity. New Phytologist. 207:437--453.

\leavevmode\hypertarget{ref-Michonneau2016}{}%
Michonneau F., Brown J.W., Winter D.J. 2016. rotl: an R package to interact with the Open Tree of Life data. Methods in Ecology and Evolution. 7:1476--1481.

\leavevmode\hypertarget{ref-morell1996roots}{}%
Morell V. 1996. The roots of phylogeny. Science. 273:569.

\leavevmode\hypertarget{ref-Morlon2014}{}%
Morlon H. 2014. Phylogenetic approaches for studying diversification. Ecology Letters. 17:508--525.

\leavevmode\hypertarget{ref-Omeara2019}{}%
O'Meara B., Md Tayeen A.S., Sanchez Reyes L.L. 2019. Rphylotastic: An r interface to 'phylotastic' web services..

\leavevmode\hypertarget{ref-Ooms2018}{}%
Ooms J., Chamberlain S. 2018. Phylocomr: Interface to 'phylocom'..

\leavevmode\hypertarget{ref-Paradis2004}{}%
Paradis E., Claude J., Strimmer K. 2004. APE: analyses of phylogenetics and evolution in R language. Bioinformatics. 20:289--290.

\leavevmode\hypertarget{ref-pearse2013phylogenerator}{}%
Pearse W.D., Purvis A. 2013. PhyloGenerator: An automated phylogeny generation tool for ecologists. Methods in Ecology and Evolution. 4:692--698.

\leavevmode\hypertarget{ref-Piel2002}{}%
Piel W.H., Donoghue M., Sanderson M. 2002. TreeBASE : A database of phylogenetic information. In: Shimura J., Wilson K., Gordon D., editors. To the interoperable ``catalog of life'' with partners. Tsukuba, Japan: National Institute for Environmental Studies. p. 41--47.

\leavevmode\hypertarget{ref-posadas2006historical}{}%
Posadas P., Crisci J.V., Katinas L. 2006. Historical biogeography: A review of its basic concepts and critical issues. Journal of Arid Environments. 66:389--403.

\leavevmode\hypertarget{ref-price2014niche}{}%
Price T.D., Hooper D.M., Buchanan C.D., Johansson U.S., Tietze D.T., Alström P., Olsson U., Ghosh-Harihar M., Ishtiaq F., Gupta S.K., others. 2014. Niche filling slows the diversification of himalayan songbirds. Nature. 509:222.

\leavevmode\hypertarget{ref-RCoreTeam2018}{}%
R Core Team. 2018. R: a language and environment for statistical computing. Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Revell2012}{}%
Revell L.J. 2012. Phytools: An r package for phylogenetic comparative biology (and other things). Methods in Ecology and Evolution. 3:217--223.

\leavevmode\hypertarget{ref-Ronquist2003}{}%
Ronquist F., Huelsenbeck J.P. 2003. MrBayes 3: Bayesian phylogenetic inference under mixed models. Bioinformatics. 19:1572--1574.

\leavevmode\hypertarget{ref-Schenk2016}{}%
Schenk J.J. 2016. Consequences of secondary calibrations on divergence time estimates. PLoS ONE. 11.

\leavevmode\hypertarget{ref-Smith2012}{}%
Smith S.A., O'Meara B.C. 2012. TreePL: Divergence time estimation using penalized likelihood for large phylogenies. Bioinformatics. 28:2689--2690.

\leavevmode\hypertarget{ref-Stoltzfus2013}{}%
Stoltzfus A., Lapp H., Matasci N., Deus H., Sidlauskas B., Zmasek C.M., Vaidya G., Pontelli E., Cranston K., Vos R., Webb C.O., Harmon L.J., Pirrung M., O'Meara B., Pennell M.W., Mirarab S., Rosenberg M.S., Balhoff J.P., Bik H.M., Heath T.A., Midford P.E., Brown J.W., McTavish E.J., Sukumaran J., Westneat M., Alfaro M.E., Steele A., Jordan G. 2013. Phylotastic! Making tree-of-life knowledge accessible, reusable and convenient. BMC Bioinformatics. 14.

\leavevmode\hypertarget{ref-title2016macrophylogenies}{}%
Title P.O., Rabosky D.L. 2016. Do Macrophylogenies Yield Stable Macroevolutionary Inferences? An Example from Squamate Reptiles. Systematic Biology.:syw102.

\leavevmode\hypertarget{ref-Webb2000}{}%
Webb C.O. 2000. Exploring the Phylogenetic Structure of Ecological Communities : An Example for Rain Forest Trees. The American Naturalist. 156:145--155.

\leavevmode\hypertarget{ref-Webb2008}{}%
Webb C.O., Ackerly D.D., Kembel S.W. 2008. Phylocom: Software for the analysis of phylogenetic community structure and trait evolution. Bioinformatics. 24:2098--2100.

\leavevmode\hypertarget{ref-webb2005phylomatic}{}%
Webb C.O., Donoghue M.J. 2005. Phylomatic: Tree assembly for applied phylogenetics. Molecular Ecology Notes. 5:181--183.

\newpage

\begin{center}
\textsc{Figure \ref{fig:workflow}}
\end{center}
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with \texttt{datelife}, via the R package or through the website. Details on the functions involved on each workflow are shown in \texttt{datelife}'s R package vignette.

\begin{center}
\textsc{Figure \ref{fig:runtime1}}
\end{center}
Computation time of query processing and search across \texttt{datelife}'s chronogram database relative to number of input taxon names. We sampled N names from the class Aves for each cohort 100 times and then performed a search with query processing not using the Taxon Names Resoultion Service (TNRS; dark gray), and using TNRS (light gray). We also performed a search using the already processed query for comparison (light blue).

\begin{center}
\textsc{Figure \ref{fig:schronograms1}}
\end{center}
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: Barker et al. (\protect\hyperlink{ref-barker2012going}{2012}), 2: Barker et al. (\protect\hyperlink{ref-barker2015new}{2015}), 3: Burns et al. (\protect\hyperlink{ref-burns2014phylogenetics}{2014}), 4: Claramunt and Cracraft (\protect\hyperlink{ref-claramunt2015new}{2015}), 5: Gibb et al. (\protect\hyperlink{ref-gibb2015new}{2015}), 6: Hedges et al. (\protect\hyperlink{ref-Hedges2015}{2015}), 7: Hooper and Price (\protect\hyperlink{ref-hooper2017chromosomal}{2017}), 8: Jetz et al. (\protect\hyperlink{ref-Jetz2012}{2012}), 9: Price et al. (\protect\hyperlink{ref-price2014niche}{2014}).

\begin{center}
\textsc{Figure \ref{fig:summaries}}
\end{center}
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarising information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

\begin{center}
\textsc{Figure \ref{fig:cvbladj}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees without branch lengths dated using BLADJ. The dating analysis can only be performed in trees with more than 2 tips, thus excluding chronogram from study 4; its data was still used as calibration for the other source chronograms.

\begin{center}
\textsc{Figure \ref{fig:cvbold}}
\end{center}
LTT plots showing results from the cross-validation analyses of trees with branch length reconstructed with data from the Barcode of Life Database (BOLD) dated using PATHd8. We could construct a tree with branch lengths for all source chronograms. However, dating with PATHd8 was only successful in three source chronograms shown here.

\newpage

\begin{figure}[!h]
\includegraphics{Fig1.pdf}
\caption{}
\label{fig:workflow}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics[width=1\linewidth]{fig_runtime1.pdf}
\caption{}
\label{fig:runtime1}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics[width=1\linewidth]{fig_schronograms1.pdf}
\caption{}
\label{fig:schronograms1}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_summaries.pdf}
%\includegraphics{fig_summaries.pdf}
\caption{}
\label{fig:summaries}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_crossval_bladj.pdf}
\caption{}
\label{fig:cvbladj}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_crossval_boldsumm.pdf}
\caption{}
\label{fig:cvbold}
\end{figure}

\end{document}