### 2019-06-23 table0_pdf_template.Rmd
birds plan

---
title: "Datelife Reports"
author: "Luna L. Sanchez Reyes"
date: "`r Sys.Date()`"
output: bookdown::pdf_document2
bibliography: library_red.bib
csl: systematic-biology.csl
link-citations: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{makecell}
- \usepackage{multirow}
- \usepackage{longtable}
- \usepackage{caption}
- \usepackage{array}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{xcolor}
geometry: "left=2.5cm,right=2.5cm,top=2cm,bottom=2cm"
vignette: >
  %\VignetteIndexEntry{DateLife Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---
```{r setup, include = FALSE}
# print(getwd())
# source("report_template_setup.R")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
my_table_format <- function(tt){
  res <- format.data.frame(tt, digits = 4, nsmall = 4, justify = "none")
  return(res)
}
make_table0 <- function(tax_dr, tax_summ, tax_dq, table_caption = "", label = "source_chr"){
    # to determine taxon number, take only the first chronogram from the same publication
    # the most common case is that all chronograms in the same publication have the same species sampling
    # but there are cases in which this is not the case (e.g., De-Nova et al. 2018)
    # and there are several chronograms in the same publication that have different taxon sampling
    # a way to incorporate them would be putting the taxon presence of all source chronograms from the same publication in the same vector, separated by "\n"
    # for now, we will just show the maximum number of sampled species in study
    nn <- sort(unique(names(tax_dr)))
    # taxon_number_all <- sapply(seq(nrow(tax_summ$matrix)), function(x) sum(tax_summ$matrix[x,]))
    # the line above is replaced by tax_summ$summary2$taxon_number
    taxon_number <- unname(sapply(nn, function(x) max(tax_summ$summary2$taxon_number[names(tax_dr) %in% x])))
    xx <- table(names(tax_dr))
    table0 <- data.frame(var1 = paste0(seq(length(xx)), "."),
        var2 = names(xx), var3 = as.numeric(unname(xx)),
        var4 = paste0(taxon_number, "/", length(tax_dq$cleaned_names)))
    table0 <- dplyr::mutate(table0, var1 = kableExtra::cell_spec(var1, "latex", align = "r",
        bold = TRUE, italic = TRUE),
        var2 = kableExtra::cell_spec(var2, "latex", font_size = 8),
        var3 = kableExtra::cell_spec(var3, "latex", align = "c"),
        var4 = kableExtra::cell_spec(var4, "latex", align = "c"))
    t0 <- knitr::kable(table0, escape = FALSE, row.names = FALSE, format = "latex",
        booktabs = TRUE, caption = table_caption, label = label, linesep = "", longtable = TRUE)
    t0 <- gsub("var1", "   ", t0)
    t0 <- gsub("var2", "Citation", t0)
    t0 <- gsub("var3", "Source N", t0)
    t0 <- gsub("var4", "Taxon N", t0)
    kableExtra::kable_styling(kable_input = t0, latex_options = "hold_position") %>%
    column_spec(1, width = "0.4cm") %>%
    column_spec(2, width = "11cm") %>%
    column_spec(3, width = "1.5cm") %>%
    column_spec(4, width = "1.8cm") %>%
    # collapse_rows(columns = 3, latex_hline = "none", valign = "middle") %>%
    row_spec(0, bold = TRUE, align = "c", italic = TRUE, font_size = 9) %>%
    footnote(general = c("Source_N: Number of source chronograms reported in study.",
            "Taxon_N: Number of queried taxa found in source chronograms."),
            general_title = "") -> t0
    t0 <- gsub("Source\\\\_N", "{\\\\textbf{\\\\textit{Source N}}}", t0)
    t0 <- gsub("Taxon\\\\_N", "{\\\\textbf{\\\\textit{Taxon N}}}", t0)
    return(t0)
}
library(magrittr)
loadd(tol_dq)
loadd(tol_dr)
loadd(tol_summ)
loadd(table0_caption)
loadd(taxon)
```

# Datelife result: `r taxon`

```{r echo = FALSE, warning=FALSE, message = FALSE, results = 'asis'}
make_table0(tol_dr, tol_summ, tol_dq, table0_caption)
```