### 2019-09-30 biorxiv_2019.09.28.Rmd
for reproducibility's sake

---
output:
  bookdown::pdf_document2:
    keep_tex: yes
    toc: false
    includes:
      # in_header: preamble.tex
      after_body: [biorxiv_2019.09.28figcaps.tex]
    # fig_caption: no
textalign: left  # this does not work...
geometry: margin = 1in
bibliography: paper_references.bib
csl: systematic-biology.csl
link-citations: yes
# indent: true # this works
# header-includes:
#    - \setlength\parindent{24pt}  # this works too, but I'm gonna indent manually with 4 spaces
# nope, 4 spaces indents are not recognized, they do something else
linestretch: 2
header-includes:
    - \usepackage{caption}
    - \usepackage{blindtext}
    - \usepackage[left]{lineno}
    - \linenumbers
  # - \usepackage{xr}
    # - \usepackage[utf8]{inputenc}
    # - \usepackage[english]{babel}
    # - \usepackage[document]{ragged2e}
nocite: |
  @barker2012going, @barker2015new, @burns2014phylogenetics, @claramunt2015new, @gibb2015new, @Hedges2015, @hooper2017chromosomal, @Jetz2012, @price2014niche
---
<!-- \begin{\nolinenumbers} -->
<!-- Running head: DATELIFE: REVEALING THE DATED TREE OF LIFE <!--MAX 50 CHARACTERS, ALL CAPITALS-->

**Title**: DateLife: Leveraging databases and analytical tools to reveal the dated Tree of Life

**Authors**: Luna L. Sánchez-Reyes^1^, Brian C. O’Meara^1^

**Correspondence address**:

1. *Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville, 425 Hesler Biology Building, Knoxville, TN 37996, USA*

**Corresponding authors**: sanchez.reyes.luna@gmail.com, bomeara@utk.edu
<!-- \end{nolinenumbers} -->

``` {r child = 'paper.Rmd'}
```
\newpage
\begin{center}
\textsc{References}
\end{center}