### 2019-09-30 paper_drafts/bioarxiv_2019.09/paper_figures.Rmd

Figure 1. 
Stylized DateLife workflow. This shows the general worflows and analyses that can be performed with DateLife, via the R package or through the website. Details on the functions involved on each workflow are shown in `datelife`'s R package vignette.

Figure 2. 
Computation time of input processing and search across `datelife`s chronogram database.

Figure 3. 
Lineage through time (LTT) plots of source chronograms containing all or a subset of species from the bird family Fringillidae of true finches. Arrows indicate maximum age of each chronogram. Numbers reference to chronograms' original publications 1: @barker2012going, 2: @barker2015new, 3: @burns2014phylogenetics, 4: @claramunt2015new, 5: @gibb2015new, 6: @Hedges2015, 7: @hooper2017chromosomal, 8: @Jetz2012, 9: @price2014niche.

Figure 4.
LTT plots of median and Supermatrix Distance Method (SDM) chronograms summarizing information from source chronograms found for the Fringillidae. Arrows indicate maximum age.

\newpage

\begin{figure}[!h]
\includegraphics{Fig1.pdf}
\caption{}
\label{fig:workflow}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_runtime1.pdf}
\caption{}
\label{fig:runtime1}
\end{figure}

\newpage

\begin{figure}[!h]
\includegraphics{fig_schronograms1.pdf}
\caption{}
\label{fig:schronograms}
\end{figure}

\newpage

\begin{figure}[!ht]
\includegraphics{fig_crossval_bladj.pdf}
\caption{}
\label{fig:cvbladj}
\end{figure}