### 2020-04-25 paper_1-header.Rmd

---
output:
  pdf_document: default
  html_document: default
---

**Title**: DateLife: Leveraging databases and analytical tools to reveal the dated Tree of Life

**Authors**: Luna L. Sánchez-Reyes^1^, Brian C. O’Meara^1^

**Correspondence address**:

1. *Department of Ecology and Evolutionary Biology, University of Tennessee, Knoxville, 425 Hesler Biology Building, Knoxville, TN 37996, USA*

**Corresponding authors**: sanchez.reyes.luna@gmail.com, bomeara@utk.edu
<!-- \end{nolinenumbers} -->